---
title: "P8130_hw5_jsg2145"
author: "Jared Garfinkel"
date: "12/6/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(faraway)
library(arsenal)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
``````

```{r}
states <- attributes(state.x77) %>% 
  .[[2]] %>% 
  .[[1]]

df <- cbind(states, state.x77) %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(population = as.numeric(population),
         income = as.numeric(income),
         illiteracy = as.numeric(illiteracy),
         life_exp = as.numeric(life_exp),
         murder = as.numeric(murder),
         hs_grad = as.numeric(hs_grad),
         frost = as.numeric(frost),
         area = as.numeric(area))
```

```{r}
tab1 <- tableby( ~ population + income + illiteracy + life_exp + murder + hs_grad + frost + area, data=df)

my_labels <- list(population = "Population", income = "Per Capita Income", illiteracy = "Illiteracy, Percent of Population", life_exp = "Life Expectancy in Years", murder = "Murder Rate per 100,000", hs_grad = "Percent High-School Graduates", frost = "Mean Number of Days with Minimum Temp Below Freezing", area = "Land Area (sq mi)")



```

