---
title: "P8130_hw5_jsg2145"
author: "Jared Garfinkel"
date: "12/6/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(faraway)
library(arsenal)
library(leaps)
library(caret)
library(broom)
library(modelr)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
``````

```{r}
states <- attributes(state.x77) %>% 
  .[[2]] %>% 
  .[[1]]

df <- cbind(states, state.x77) %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(population = as.numeric(population),
         income = as.numeric(income),
         illiteracy = as.numeric(illiteracy),
         life_exp = as.numeric(life_exp),
         murder = as.numeric(murder),
         hs_grad = as.numeric(hs_grad),
         frost = as.numeric(frost),
         area = as.numeric(area))

str(df)
```

```{r}
tab1 <- tableby( ~ population + income + illiteracy + life_exp + murder + hs_grad + frost + area, data=df)

my_labels <- list(population = "Population", 
                  income = "Per Capita Income", 
                  illiteracy = "Illiteracy, Percent of Population", 
                  life_exp = "Life Expectancy in Years", 
                  murder = "Murder Rate per 100,000", 
                  hs_grad = "Percent High-School Graduates", 
                  frost = "Mean Number of Days with Minimum Temp Below Freezing", 
                  area = "Land Area (sq mi)")

my_controls <- tableby.control(
               total = T,
               test = F,  # No test p-values yet
               numeric.stats = c("meansd", "medianq1q3", "range"),
               stats.labels = list(
               meansd = "Mean (SD)",
               medianq1q3 = "Median (Q1, Q3)",
               range = "Min - Max"))

tab2 <- tableby( ~ population + income + illiteracy + life_exp + murder + hs_grad + frost + area, data=df, control=my_controls)

summary(tab2, title = "Descriptive Statistics: Life Expectancy Data", labelTranslations = my_labels, text=T) %>% 
  data.frame() %>% 
  knitr::kable()

```

```{r}
df %>% 
  ggplot(aes(x = life_exp)) + 
  geom_histogram(bins = 10)
```

```{r}
skimr::skim(df)
```

```{r}
df %>% 
  ggplot(aes(x = frost)) +
  geom_histogram(bins = 10)
```

 
```{r}
df %>% 
  ggplot(aes(x = log(population))) +
  geom_histogram(bins = 10)
```

```{r}
df %>% 
  ggplot(aes(x = income, y = life_exp)) +
  geom_point()
```

```{r}
df %>% 
  ggplot(aes(x = illiteracy, y = life_exp)) +
  geom_point()
```

```{r}
df <- df %>% 
  mutate(ln_pop = log(population),
         ln_area = log(area))
```

```{r}
fit1 <- lm(life_exp ~ population, data = df)
broom::tidy(fit1)$p.value[2]
fit2 <- lm(life_exp ~ murder, data = df)
broom::tidy(fit2)$p.value[2]
fit3 <- lm(life_exp ~ income, data = df)
broom::tidy(fit3)$p.value[2]
fit4 <- lm(life_exp ~ illiteracy, data = df)
broom::tidy(fit4)$p.value[2]
fit5 <- lm(life_exp ~ hs_grad, data = df)
broom::tidy(fit5)$p.value[2]
fit6 <- lm(life_exp ~ frost, data =df)
broom::tidy(fit6)$p.value[2]
fit7 <- lm(life_exp ~ area, data = df)
broom::tidy(fit7)$p.value[2]
fit8 <- lm(life_exp ~ log(population), data = df)
broom::tidy(fit8)$p.value[2]
fit9 <- lm(life_exp ~ log(area), data = df)
broom::tidy(fit9)$p.value[2]
```

```{r}
# murder has the lowest p-value

forward1<-lm(life_exp ~ murder, data = df)
broom::tidy(forward1)

fit1 <- update(forward1, . ~ . + income)
broom::tidy(fit1)$p.value[3]

fit2 <- update(forward1, . ~ . + ln_pop)
broom::tidy(fit2)$p.value[3]

fit3 <- update(forward1, . ~ . + illiteracy)
broom::tidy(fit3)$p.value[3]

fit4 <- update(forward1, . ~ . + hs_grad)
broom::tidy(fit4)$p.value[3] 

fit5 <- update(forward1, . ~ . + frost)
broom::tidy(fit5)$p.value[3] 

fit6 <- update(forward1, . ~ . + ln_area)
broom::tidy(fit6)$p.value[3] 
```

```{r}
# next lowest p-value is hs_grad

forward2 <- update(forward1, . ~ . + hs_grad)
tidy(forward2)

fit1 <- update(forward2, . ~ . + ln_pop)
tidy(fit1)$p.value[4] 

fit2 <- update(forward2, . ~ . + income)
tidy(fit2)$p.value[4] 

fit3 <- update(forward2, . ~ . + illiteracy)
tidy(fit3)$p.value[4] 

fit4 <- update(forward2, . ~ . + frost)
tidy(fit4)$p.value[4] 

fit5 <- update(forward2, . ~ . + ln_area)
tidy(fit5)$p.value[4] 
```

```{r}
# next lowest p-value is the log of the population

forward3 = update(forward2, . ~ . + ln_pop)
tidy(forward3)

fit1 <- update(forward3, . ~ . + income)
tidy(fit1)$p.value[5]

fit2 <- update(forward3, . ~ . + illiteracy)
tidy(fit2)$p.value[5]

fit3 <- update(forward3, . ~ . + frost)
tidy(fit3)$p.value[5]

fit4 <- update(forward3, . ~ . + ln_area)
tidy(fit4)$p.value[5]
```

```{r}
# next lowest p-value is frost

forward4 = update(forward3, . ~ . + frost)
tidy(forward4)

fit1 <- update(forward4, . ~ . + income)
tidy(fit1)$p.value[6] # 0.8868104

fit2 <- update(forward4, . ~ . + illiteracy)
tidy(fit2)$p.value[6] # 0.7542022

fit3 <- update(forward4, . ~ . + ln_area)
tidy(fit3)$p.value[6] # 0.5212581
```

```{r}
fit_back <- lm(life_exp ~ . -states -population -area, data = df)
broom::tidy(fit_back) ## remove income
fit_back2 <- lm(life_exp ~ . -states -population -area -income, data = df)
broom::tidy(fit_back2) ## remove illiteracy
fit_back3 <- lm(life_exp ~ . -states -population -area -income -illiteracy, data = df)
broom::tidy(fit_back3) ## remove ln_area
fit_back3 <- lm(life_exp ~ . -states -population -area -income -illiteracy -ln_area, data = df)
broom::tidy(fit_back3) ## remove

vif(fit_back)
summary(fit_back3)

fit_states <- lm(life_exp ~ states, data = df)
anova(fit_states)
```

```{r}
df %>% 
  ggplot(aes(x = illiteracy, y = hs_grad)) +
  geom_point()
```

```{r}
fit <- lm(life_exp ~ hs_grad + illiteracy, data = df)
vif(fit)
summary(fit)

```



```{r}
model1 <- step(fit1, direction = "both")
model2 <- step(fit1, direction = "backward")
model3 <- step(fit_back, direction = "forward")
model2 <- step
```

```{r}
output = vector(mode = "list", length = 7)

df = df %>% 
  select(-states)

for (i in 1:7) {
  output[[i]] = lm(life_exp ~ df[[i]], data = df) %>%
    broom::tidy() %>% 
    tibble(
      variable = colnames(df[i]),
      pvalue = .$p.value[[2]]
    ) %>% 
    select(variable, pvalue) %>% 
    bind_rows()
}
```

```{r}
df = df %>% 
  select(life_exp, everything(), -population, -area) %>% 
  data.frame()

leaps(x = df[,2:8], y = df[,1], nbest = 2, method = "Cp")

leaps(x = df[,2:8], y = df[,1], nbest = 2, method = "adjr2")
```

```{r}
b <- regsubsets(life_exp ~ ., data=df)
   (rs <- summary(b))
```


```{r}
par(mar=c(4,4,1,1))
par(mfrow=c(1,2))

plot(2:8, rs$cp, xlab="No of parameters", ylab="Cp Statistic")
abline(0,1)

plot(2:8, rs$adjr2, xlab="No of parameters", ylab="Adj R2")
```

```{r}
# AIC of the 6-predictor model:
multi.fit4 <- lm(life_exp ~ murder + hs_grad + ln_pop + frost, data=df)
AIC(multi.fit4)

# BIC
AIC(multi.fit4, k = log(length(pull(df, life_exp))))

# AIC of the 4-predictor model:
multi.fit5 <- lm(life_exp ~ murder + hs_grad + ln_pop + frost + ln_area, data = df)
AIC(multi.fit5)

# 4-predictor model is the best
```

```{r}
par(mfrow =c(2, 2))
plot(multi.fit4)

```

```{r}
influence.measures(multi.fit4)
```


```{r}
df.no11 <- df[-11,]

multi.fit2 <- lm(life_exp ~ murder + hs_grad + ln_pop + frost, data = df.no11)

par(mfrow =c(2,2))
plot(multi.fit2)
```

```{r}
summary(multi.fit2)
summary(multi.fit4)
```

This is evidence that the influential point at observation 11 causes a variable, frost, to become not significant.

```{r}
set.seed(1)
data_train<-trainControl(method="cv", number=10)
```

```{r}
model_caret <- train(life_exp ~ murder + hs_grad + ln_pop + frost,
                   data=df,
                   trControl=data_train,
                   method='lm',
                   na.action=na.pass)

model_caret

model_caret$finalModel

model_caret$resample

sd(model_caret$resample$Rsquared)

full_model <- lm(life_exp ~ murder + hs_grad + ln_pop + frost, data = df)
summary(full_model)
```

Through automatic and criterion based model building procedures, it was determined that a final model including the murder rate, high school graduation rate, log of the state population, and mean number of days below freezing each year (frost days) could predict the life expectancy in a state with an adjusted r-squared of 0.77 $\pm$ 0.18. There was one leverage point that, when removed, caused one variable, frost days, to become not significant in the model. This indicates that there may be evidence that a model with three variables, murder rate, high_school graduation rate, and the log of the state population is the best model.

## Problem 2

```{r}
rent_df <- read_csv("./CommercialProperties.csv") %>% 
  janitor::clean_names() %>% 
  mutate(rental_rate = rental_rate*1000) %>% 
  rename(monthly = rental_rate)
```

```{r}
skimr::skim(rent_df)
```

```{r}
rent_df %>% 
  ggplot(aes(x = age - (mean(age)))) +
  geom_histogram()
```



```{r}
fit_rent <- lm(monthly ~ ., data = rent_df)

summary(fit_rent)
```

After fitting an initial model including all available variables, it appears that the age of the property, the taxes, and square footage are significantly associated with the outcome, monthly rent. However, the vacancy rate does not appear to be associated with the monthly rent.

```{r}
rent_df %>% 
  ggplot(aes(x = age, y = monthly)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
rent_df %>% 
  ggplot(aes(x = taxes, y = monthly)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

```{r}
rent_df %>% 
  ggplot(aes(x = sq_footage, y = monthly)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

Square footage and taxes appear to be directly related with the monthly rent of each property, while the age of the property might be slightly inversely related to the monthly rent. There appear to be mostly homes younger than five years old or older than ten years old in this data.

```{r}
fit_rent_sig <- lm(monthly ~ . -vacancy_rate, data = rent_df)

summary(fit_rent_sig)
```

```{r}
rent_df_tr <- rent_df %>% 
  mutate(age_0 = age - mean(age),
         age2 = age^2)
fit_rent_0 = lm(monthly ~ age_0 + taxes + sq_footage, data = rent_df_tr) 
fit_rent2  = lm(monthly ~ age2 + taxes + sq_footage, data = rent_df_tr)

summary(fit_rent_0)
summary(fit_rent2)
```

Centering the age of the property data using the mean does not change the adjusted r-squared, but using a higher order term does reduce the adjusted r-squared. For this reason, it may not be necessary to use one of these transformations. On the other hand, because there are appear to be two clusters of the ages of properties, it may make sense to use piecewise linear regression. We create a break at 10, as appears to be a visual break.

```{r}
rent_df_pwl = rent_df %>% 
  mutate(age_pwl5  = (age - 15) * (age >= 15))
```

```{r}
fit_pwl = lm(monthly ~ age + age_pwl5, data = rent_df_pwl)
fit_lin_age = lm(monthly ~ age, data = rent_df_pwl)

anova(fit_lin_age, fit_pwl)
```

```{r}
summary(fit_pwl)
```

```{r}
rent_df_pwl %>% 
  gather_predictions(fit_lin_age, fit_pwl) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = age, y = monthly)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

A piecewise linear model does not seem to improve the association between age and monthly rent. So, a model including all the significant variables is the recommended model for the relationship between monthly rent and other variables.